---
title: "Tutorial: Merging of Final Output Data Sets"
author: "André Vidas Olsen"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Introduction
This tutorial will cover how to merge multiple final output files into one file.


#Data
The data for this tutorial will consist of the following files:

* The final output files generated from multiple LipidQ quantification runs
(see quantification tutorial for info on how to use the quantification procedure
in the LipidQ GUI).


#Step 1: open LipidQ
Start by open up LipidQ. This can be done by doing these 3 steps:

1. Open Rstudio

2. Write "library(lipidQ)"

3. Write "runLipidQ()"

If the web browser encounters problems doing any of the steps in the tutorial,
LipidQ can also be open by going to:
```{r, echo=FALSE}
system.file("app/", package = "lipidQ")
```
, open either ui.R or server.R and hit the Run button at the right top corner of
Rstudio. 

#Step 2. Load data into lipidQ
After opening LipidQ, the operation tab will start at the Quantification tab.
Instead of staying at this operation tab, click on the "Merging of Final Output
Data Sets". In this section, the first field consists of choosing the input
data, which in this case is the output files generated from multiple
quantification runs with LipidQ. Use the "Browse..." button to browse and
select the desired input data for merging process. 

#Step 3. Select ouput folder
After selecting the data to be merged, you need to specify the output folder.
This is done by writing or copy-pasting the path to the desired folder in the
field below "Paste the filepath for the output files to be saved". The following
describes how to obtain the full path:

From Mac OS X El Capitan 2015 onwards (As of February 2018): Go to the folder of
interest, right click on the folder, press the alt button + click on 'Copy
"folder-name" as Pathname'. Go back to LipidQ and paste the folder path
(&#8984; Command button + V) into the output folder field.

Windows: Go to the folder of interest, right click on the folder, go to
properties, find Location and copy the path there (Ctrl + C). Go back to
LipidQ and paste the full path (Ctrl + V) into the output folder field.
Remember to change every "\\" in the path to "/".


After the folder path has been written, click on the "Merge Final Output Data
Sets" to start the process.
