---
title: "Tutorial: Merging of Final Output Data Sets"
author: "André Vidas Olsen"
date: "`r Sys.Date()`"
output: html_document
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

#Introduction
This tutorial will cover how to merge multiple final output files into one file.


#Data
The data for this tutorial will consist of the following files:

* The final output files generated from multiple LipiqQuan quantification runs (see quantification tutorial for info on how to use the quantification procedure in the LipidQuan GUI).


#Step 1: open LipidQuan
Start by open up LipidQuan. This can be done by doing these 3 steps:

1. Open Rstudio

2. Write "library(lipidQuan)"

3. Write "runLipidQuan()"

If the web browser encounters problems doing any of the steps in the tutorial, LipidQuan can also be open by going to:
```{r, echo=FALSE}
system.file("app/", package = "lipidQuan")
```
, open either ui.R or server.R and hit the Run button at the right top corner of Rstudio. 

#Step 2. Load data into lipidQuan
After opening LipidQuan, the operation tab will start at the Quantification tab. Instead of staying at this operation tab, click on the "Merging of Final Output Data Sets". In this section, the first field consists of choosing the input data, which in this case is the output files generated from multiple quantification runs with LipidQuan. Use the "Browse..." button to browse and select the desired input data for merging process. 

#Step 3. Select ouput folder
After selecting the data to be merged, you need to specify the output folder. This is done by writing or copy-pasting the path to the desired folder in the field below "Paste the filepath for the output files to be saved". After the folder path has been written, click on the "Merge Final Output Data Sets" to start the process.
